---
title: "Morphometricity analysis by ROI"
author: "by [Baptiste Couvy-Duchesne] - `r format(Sys.time(), '%d %B %Y')`"
output:
  epuRate::PCTG:
    toc: TRUE
    code_folding: "show"
---

# Create list of vertices for each cortical and subcortical regions, for left and right hemispheres

See *atlas* folder for the **fsaverage_vertex_labels_names_order2.txt** file 

**You do not need to run this first section, unless you want to use a different atlas.** Here we used cortical regions defined by the Desikan atlas. We extracted from FreeSurfer the information relative to each vertex coordinates (may be useful for plotting) and ROI they belong to.  

## Cortical ROIs

```{R, message=FALSE, eval=F}
label=read.table("atlas/fsaverage_vertex_labels_names_order2.txt", header=T, stringsAsFactors = F)
label$RHLabel[which(is.na(label$RHLabel))]<-"unknown"
label$LHLabel[which(is.na(label$LHLabel))]<-"unknown"

range(table(label$RHLabel)) # ROI vary in term or number of vertices
range(table(label$LHLabel)) # ROI vary in term or number of vertices

# Modify vertex ID to match names in bod files
label$VertexID=substr(label$VertexNum, 11, nchar(label$VertexNum))

# LEFT
for (roi in unique(label$LHLabel)){
  vertices2Write=label$VertexID[which(label$LHLabel==roi)]
write.table(paste0("lha_",vertices2Write), paste0("atlas/Vertices_", roi,"_lh_area_UKB.txt"), col.names = F, row.names = F, quote=F )
write.table(paste0("lht_",vertices2Write), paste0("atlas/Vertices_", roi,"_lh_thickness_UKB.txt"), col.names = F, row.names = F, quote=F )
}

# RIGHT
for (roi in unique(label$RHLabel)){
  vertices2Write=label$VertexID[which(label$RHLabel==roi)]
write.table(paste0("rha_",vertices2Write), paste0("atlas/Vertices_", roi,"_rh_area_UKB.txt"), col.names = F, row.names = F, quote=F )
write.table(paste0("rht_",vertices2Write), paste0("atlas/Vertices_", roi,"_rh_thickness_UKB.txt"), col.names = F, row.names = F, quote=F )
}
```

## Subcortical ROIs

```{R, message=FALSE, eval=F}

label=read.table("atlas/fsaverage_subcortical_vertices_label.txt", header=T, stringsAsFactors = F)
range(table(label$subcvName))

ROIS=c("Thalamus-Proper",  "Caudate", "Putamen","Pallidum","Hippocampus","Amygdala","Accumbens-area")

# LEFT
for (roi in ROIS){
  vertices2Write=label$ProbeID[which(label$subcvName==paste0("Left-", roi))]
write.table(paste0("LogJacs_",vertices2Write), paste0("atlas/Vertices_", roi,"_lh_LogJacs_UKB.txt"), col.names = F, row.names = F, quote=F )
write.table(paste0("thick_",vertices2Write), paste0("atlas/Vertices_", roi,"_lh_thick_UKB.txt"), col.names = F, row.names = F, quote=F )
}

# RIGHT
for (roi in ROIS){
  vertices2Write=label$ProbeID[which(label$subcvName==paste0("Right-", roi))]
write.table(paste0("LogJacs_",vertices2Write), paste0("atlas/Vertices_", roi,"_rh_LogJacs_UKB.txt"), col.names = F, row.names = F, quote=F )
write.table(paste0("thick_",vertices2Write), paste0("atlas/Vertices_", roi,"_rh_thick_UKB.txt"), col.names = F, row.names = F, quote=F )
}

```

# Create BRMs corresponding to each cortical and subcortical region

## Cortical

```{bash, message=FALSE, eval=FALSE}

# Integrate loops on hemisphere and modality inside each job to limit the number of jobs submitted

for ROI in lingual isthmuscingulate pericalcarine precuneus unknown postcentral precentral parsopercularis supramarginal superiorfrontal caudalmiddlefrontal bankssts inferiorparietal middletemporal lateraloccipital fusiform temporalpole inferiortemporal entorhinal parstriangularis rostralmiddlefrontal parsorbitalis cuneus superiorparietal parahippocampal rostralanteriorcingulate medialorbitofrontal lateralorbitofrontal superiortemporal insula caudalanteriorcingulate frontalpole posteriorcingulate paracentral transversetemporal 
do 
${bind}/qsubshcom " for hemi in lh rh |;
do |;
for moda in thickness area |;
do |;
"${bind}"/osca --befile "${wd}"/BodFiles/\${hemi}.\${moda}.fwhm0.UKB.vertexQC --extract-probe "${wd}"/atlas/Vertices_"${ROI}"_\${hemi}_\${moda}_UKB.txt --make-orm-bin --out "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM |;
done |;
done |; " 1 10G BRM_${ROI} 10:00:00 "-acct=UQ-IMB-CNSG" 
done
```

## Subcortical

```{bash, message=FALSE, eval=FALSE}

for ROI in Thalamus-Proper Caudate Putamen Pallidum Hippocampus Amygdala Accumbens-area 
do 
${bind}/qsubshcom " for hemi in lh rh |;
do |;
for moda in thick LogJacs |;
do |;
"${bind}"/osca --befile "${wd}"/BodFiles/\${hemi}.\${moda}.fwhm0.UKB.vertexQC --extract-probe "${wd}"/atlas/Vertices_"${ROI}"_\${hemi}_\${moda}_UKB.txt --make-orm-bin --out "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM |;
done |;
done |; " 1 10G BRM_${ROI} 10:00:00 "-acct=UQ-IMB-CNSG" 
done

```

# Extract values for BRM QC - enforce cut-off of 5SD

## Start interactive job and open R

```{bash, message=FALSE, eval=F}

qsub -I -X -l select=1:ncpus=1:mem=20GB -l walltime=05:00:00 -A UQ-IMB-CNSG

wd="working/directory"
cd $wd
module load R
R

```

## Extract BRM  cut-off values

```{r, message=FALSE, eval=F}

source("RR_4.0_BRM_QC_functions.R")

# CORTICAL
for (ROI in c("lingual", "isthmuscingulate", "pericalcarine", "precuneus", "postcentral", "precentral", "parsopercularis", "supramarginal", "superiorfrontal", "caudalmiddlefrontal", "bankssts", "inferiorparietal", "middletemporal", "lateraloccipital", "fusiform", "temporalpole", "inferiortemporal", "entorhinal", "parstriangularis", "rostralmiddlefrontal", "parsorbitalis", "cuneus", "superiorparietal", "parahippocampal", "rostralanteriorcingulate", "medialorbitofrontal", "lateralorbitofrontal", "superiortemporal", "insula", "caudalanteriorcingulate", "frontalpole", "posteriorcingulate", "paracentral", "transversetemporal")){
 
  vals=NULL
  for (hemi in c("lh", "rh")){
   for (moda in c("thickness", "area")){
    
brm=asBRM(ReadORMBin(paste0(hemi, "_", moda, "_", ROI, "_UKB_BRM")))
vals=rbind(vals, c(mean(diag(brm)),var(diag(brm)), mean(brm[upper.tri(brm)]),var(brm[upper.tri(brm)]) , sqrt(var(brm[upper.tri(brm)]))*5 )  )
 
  } }
rownames(vals)=c(paste(rep("lh", 2), c("thickness", "area")) , paste(rep("rh", 2), c("thickness", "area")) )
colnames(vals)=c("Mean - diag", "Var - diag", "Mean - off diag", "Var - off diag", "BRM cutoff")
print(ROI)
print(vals)
write.table(vals[,5], paste0(ROI, "_BRM_cutoffs.txt"), sep="\t", col.names = F, row.names = F )
}

# SUBCORTICAL
for (ROI in c("Thalamus-Proper", "Caudate", "Putamen", "Pallidum", "Hippocampus", "Amygdala", "Accumbens-area")){
 
  vals=NULL
  for (hemi in c("lh", "rh")){
   for (moda in c("thick", "LogJacs")){
    
  brm=asBRM(ReadORMBin(paste0(hemi, "_", moda, "_", ROI, "_UKB_BRM")))
  vals=rbind(vals, c(mean(diag(brm)),var(diag(brm)), mean(brm[upper.tri(brm)]),var(brm[upper.tri(brm)]) , sqrt(var(brm[upper.tri(brm)]))*5 )  )
 
}
}
rownames(vals)=c(paste(rep("lh", 2), c("thickness", "area")) , paste(rep("rh", 2), c("thickness", "area")) )
colnames(vals)=c("Mean - diag", "Var - diag", "Mean - off diag", "Var - off diag", "BRM cutoff")
print(ROI)
print(vals)
write.table(vals[,5], paste0(ROI, "_BRM_cutoffs.txt"), sep="\t", col.names = F, row.names = F )
}

```

# QC BRMs based on the files created

## Cortical 

```{bash, message=FALSE, eval=FALSE}

for ROI in lingual isthmuscingulate pericalcarine precuneus unknown postcentral precentral parsopercularis supramarginal superiorfrontal caudalmiddlefrontal bankssts inferiorparietal middletemporal lateraloccipital fusiform temporalpole inferiortemporal entorhinal parstriangularis rostralmiddlefrontal parsorbitalis cuneus superiorparietal parahippocampal rostralanteriorcingulate medialorbitofrontal lateralorbitofrontal superiortemporal insula caudalanteriorcingulate frontalpole posteriorcingulate paracentral transversetemporal
do 
${bind}/qsubshcom " nrow=1 |;
for hemi in lh rh |;
do |;
for moda in thickness area |;
do |;
cutof=\$(sed \"\${nrow}q;d\" "${ROI}"_BRM_cutoffs.txt) |;
"${bind}"/osca --reml --orm "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM --orm-cutoff \${cutof} --orm-cutoff-2sides --make-orm --out "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM_QC |;
((nrow++)) |;
echo \${nrow} |;
done |;
done |; " 1 4G BRM_QC_${ROI} 10:00:00 "-acct=UQ-IMB-CNSG" 
done
```

## Cortical 

```{bash, message=FALSE, eval=FALSE}

for ROI in Thalamus-Proper Caudate Putamen Pallidum Hippocampus Amygdala Accumbens-area
do 
${bind}/qsubshcom " nrow=1 |;
for hemi in lh rh |;
do |;
for moda in thick LogJacs |;
do |;
cutof=\$(sed \"\${nrow}q;d\" "${ROI}"_BRM_cutoffs.txt) |;
"${bind}"/osca --reml --orm "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM --orm-cutoff \${cutof} --orm-cutoff-2sides --make-orm --out "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM_QC |;
((nrow++)) |;
echo \${nrow} |;
done |;
done |; " 1 4G BRM_QC_${ROI} 1:00:00 "-acct=UQ-IMB-CNSG" 
done

```

# Estimate ROI-morphometricity or association with each ROI

The following example uses phenotypes regressed out for baseline and body size variables    

We loop on modalities (area / thickness) and hemisphere inside the job to reduce charge on cluster : npheno x nROIs  jobs instead of npheno x nROIs x 8 jobs   

**Note that we limit ourselves to phenotypes significant in the global morphometricity analysis to reduce computational cost**    

## Cortical ROI associations

```{bash, message=FALSE, eval=FALSE}

od="working/directory/BREML_ROI_Bsln_bdsz"

for ROI in lingual isthmuscingulate pericalcarine precuneus unknown postcentral precentral parsopercularis supramarginal superiorfrontal caudalmiddlefrontal bankssts inferiorparietal middletemporal lateraloccipital fusiform temporalpole inferiortemporal entorhinal parstriangularis rostralmiddlefrontal parsorbitalis cuneus superiorparietal parahippocampal rostralanteriorcingulate medialorbitofrontal lateralorbitofrontal superiortemporal insula caudalanteriorcingulate frontalpole posteriorcingulate paracentral transversetemporal
  do 
mkdir -p ${od}/${ROI}
mkdir -p ${od}/${ROI}_job
cd ${od}/${ROI}_job # avoids having too many log files with the output

${bind}/qsubshcom "echo \${TASK_ID}.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt) |;
for hemi in lh rh |;
do |;
for moda in area thickness |;
do |;
"${bind}"/osca --reml --orm "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM_QC --pheno  "${wd}"/Phenotypes15K/UKB_phenotypes_reg_Bsln_bdsz/\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt)_reg.txt --out "${od}"/"${ROI}"/NestedModel.\${hemi}.\${moda}.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt).outliers.UKB --reml-maxit 10000 |;
done |;
done " 1 4G BREML_${ROI} 5:00:00 "-array=1-43 -acct=UQ-IMB-CNSG" 
done
```

## Subcortical ROI associations

```{bash, message=FALSE, eval=FALSE}

od="working/directory/BREML_ROI_Bsln_bdsz"

for ROI in Thalamus-Proper Caudate Putamen Pallidum Hippocampus Amygdala Accumbens-area
  do 
mkdir -p ${od}/${ROI}
mkdir -p ${od}/${ROI}_job
cd ${od}/${ROI}_job # Put all log files written by qsubshcom in the same directory

${bind}/qsubshcom "echo \${TASK_ID}.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt) |;
for hemi in lh rh |;
do |;
for moda in LogJacs thick |;
do |;
"${bind}"/osca --reml --orm "${wd}"/BodFiles/\${hemi}_\${moda}_"${ROI}"_UKB_BRM_QC --pheno  "${wd}"/Phenotypes15K/UKB_phenotypes_reg_Bsln_bdsz/\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt)_reg.txt --out "${od}"/"${ROI}"/NestedModel.\${hemi}.\${moda}.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_Phenotypes_signif_Bslne_bdsz.txt).outliers.UKB --reml-maxit 10000 |;
done |;
done " 1 4G BREML_${ROI} 48:00:00 "-array=1-43 -acct=UQ-IMB-CNSG" 
done

```

# Put all results together 

## Interactive session 

```{bash, message=FALSE, eval=F}

qsub -I -X -l select=1:ncpus=1:mem=15GB -l walltime=05:00:00 -A UQ-IMB-CNSG

wd="working/directory"
cd $wd
module load R
R

```

## Loop on output files and extract values of interest

```{r, message=FALSE, eval=F}

signifUKB<-0.05/175

formatedPhe=read.csv("UKB_allvars_categories_fancylabels_withOldNames_Jan19Update.csv", stringsAsFactors = F)
signi=read.table("UKB_Phenotypes_signif_Bslne_bdsz.txt")

formatedPhe=formatedPhe[which(formatedPhe$V1 %in% signi$V1),]
colnames(formatedPhe)=c("oldVar", "category", "variable_clean", "variable")
formatedPhe=formatedPhe[order(formatedPhe$category),]

# Initialise tables
allvar<-matrix(NA, nrow = length(formatedPhe$variable), ncol=168)
rownames(allvar)<-formatedPhe$variable
allvar=as.data.frame(allvar)

allp<-matrix(NA, nrow = length(formatedPhe$variable), ncol=168)
rownames(allp)<-formatedPhe$variable
allp=as.data.frame(allp)

allse<-matrix(NA, nrow = length(formatedPhe$variable), ncol=168)
rownames(allse)<-formatedPhe$variable
allse=as.data.frame(allse)

sample="UKB" # used in naming files

colnumber=1
# Loop on all ROI
corticalROI=c("entorhinal", "parahippocampal", "temporalpole", "fusiform", "superiortemporal", "middletemporal","inferiortemporal", "transversetemporal", "bankssts","insula","lingual","pericalcarine", "cuneus","lateraloccipital","superiorfrontal", "caudalmiddlefrontal", "rostralmiddlefrontal",  "parstriangularis",  "parsorbitalis","parsopercularis","lateralorbitofrontal","medialorbitofrontal","frontalpole",  "paracentral",  "precentral", "postcentral", "supramarginal","superiorparietal", "inferiorparietal", "precuneus", "rostralanteriorcingulate","caudalanteriorcingulate", "posteriorcingulate","isthmuscingulate",  "unknown")

brainRegion=c(rep("Temporal", 10), rep("Occipital", 4), rep("Frontal", 11), rep("Parietal", 5), rep("Cingulate", 4))

for (mod in c("thickness", "area")){
for (hemi in c("lh", "rh")){
      for (ROI in corticalROI){
for (phenotype in formatedPhe$variable){
if (file.exists(paste0("BREML_ROI_Bsln_bdsz/", ROI, "/NestedModel",".",hemi,".",mod, ".", phenotype, ".outliers.", sample, ".hsq"))){
T1<-read.table(paste0("BREML_ROI_Bsln_bdsz/", ROI, "/NestedModel",".",hemi,".",mod, ".", phenotype, ".outliers.", sample, ".hsq"),fill=T, row.names = 1, stringsAsFactors = F, header=T)

allvar[phenotype,colnumber]<-as.numeric(T1[4,1])*100
allp[phenotype,colnumber]<-as.numeric(T1[9,1])
allse[phenotype,colnumber]<-as.numeric(T1[4,2])*100
} }
      colnames(allvar)[colnumber]=paste( ifelse(hemi=="lh", "Left", "Right"), mod, ROI )
   colnumber = colnumber+1
} } }

# subcortical
subcorticalROI=c("Thalamus-Proper","Caudate", "Putamen", "Pallidum", "Hippocampus", "Amygdala", "Accumbens-area")

for (mod in c("thick", "LogJacs")){
for (hemi in c("lh", "rh")){
    for (ROI in subcorticalROI){
for (phenotype in formatedPhe$variable){
if (file.exists(paste0("BREML_ROI_Bsln_bdsz/", ROI, "/NestedModel",".",hemi,".",mod, ".", phenotype, ".outliers.", sample, ".hsq"))){
T1<-read.table(paste0("BREML_ROI_Bsln_bdsz/", ROI, "/NestedModel",".",hemi,".",mod, ".", phenotype, ".outliers.", sample, ".hsq"),fill=T, row.names = 1, stringsAsFactors = F, header=T)
allvar[phenotype,colnumber]<-as.numeric(T1[4,1])*100
allp[phenotype,colnumber]<-as.numeric(T1[9,1])
allse[phenotype,colnumber]<-as.numeric(T1[4,2])*100
} }
      colnames(allvar)[colnumber]=paste( ifelse(hemi=="lh", "Left", "Right"), ifelse(mod=="thick", "thickness", "curvature" ), ROI )
   colnumber = colnumber+1
} } }

# Write tables 
write.table(allvar, "allvar_table_UKB_Bsln_bdsz_ROIs.txt", sep="\t", quote=F)
write.table(allse, "allse_table_UKB_Bsln_bdsz_ROIs.txt", sep="\t", quote=F)
write.table(allp, "allp_table_UKB_Bsln_bdsz_ROIs.txt", sep="\t", quote=F)
colSums(is.na(allvar))

```

# Plot using corrplot package

## Format tables

```{r, message=FALSE, eval=F}

# Add fixed effect R2
fixEff=read.csv("FixedEffects_UKB_15K_Jan2019Update.csv", stringsAsFactors = F)
fixEff=fixEff[,c("Variable",  "r0" , "rage", "rsex" , "rbrain", "rbody"  )]
# fixed eff for extra phenotypes
fixEff$r2=rowSums(fixEff[,c("r0" , "rage", "rsex" , "rbrain",  "rbody" )])

allvar$variable=rownames(allvar)
allROI_fixedEff=merge(allvar, fixEff[,c("Variable", "r2")], by.x="variable", by.y="Variable", sort = F)

# Correct for fixed effect R2
for (iii in 1:length(formatedPhe$variable)){
allROI_fixedEff[iii,2:169]=allROI_fixedEff[iii,2:169] * (1-allROI_fixedEff$r2[iii])
}

allROI_fixedEff$variable=NULL
allROI_fixedEff$r2=NULL

# Exclude columns of missing values and add rownames
allvar0=as.matrix(allROI_fixedEff)
rownames(allvar0)=formatedPhe$variable_clean

allp0=as.matrix(allp)
rownames(allp0)=formatedPhe$variable_clean

toExclude=which(colSums(is.na(allvar))==length(formatedPhe$variable))
toExclude
allvar0=allvar0[,-toExclude]
allp0=allp0[,-toExclude]

# Change column names
for (cn in 1:length(colnames(allvar0))){
 colnames(allvar0)[cn]= tail(strsplit(colnames(allvar0),split=" ")[[cn]],1)
}

# Significance threshold
signif_ROI=0.05/(58*176)

allvar0[which(is.na(allvar0), arr.ind = T)]=0
range(allvar0)   
    
# Set twinning results to 0 - not enough N
allvar0["Part of multiple birth",]=0
```

## Generate plots

```{r, message=FALSE, eval=F}

library(corrplot)
library(pBrackets)
library(grDevices)

png("Corplot_rB_signif_ordered_allVC_Bsln_bdsz.png", width = 80, height = 30, units = "cm", res = 300)

corrplot(allvar0, is.corr = F,  tl.cex = 1 ,  tl.col = "black",  col=colorRampPalette(c("white","white", "#F0E442","#D55E00"))(100), method = "color", p.mat = allp0, sig.level = signif_ROI, insig = "blank", na.label = "square", na.label.col = "white",  tl.srt=90, cl.ratio = 0.05, addgrid.col = "grey", cl.lim=c(0,50), cl.cex = 1.3)

vOffset=49 # This is a manual parameter to move the curly brackets at the top

# Add left right CT subcort labels
text(x = 17, y = vOffset+4, labels = "Left cortical thickness", col="grey",xpd = T)
  brackets(1, vOffset+3, 34, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
text(x = 17+35, y = vOffset+4, labels = "Right cortical thickness", col="grey",xpd = T)
  brackets(35, vOffset+3, 68, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
text(x = 17+69, y = vOffset+4, labels = "Left cortical area", col="grey",xpd = T)
  brackets(69, vOffset+3, 102, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
text(x = 17+103, y = vOffset+4, labels = "Right cortical area", col="grey",xpd = T)
  brackets(103, vOffset+3, 136, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
text(x = 137+7, y = vOffset+4, labels = "Subcortical thickness", col="grey",xpd = T)
  brackets(137, vOffset+3, 150, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
text(x = 150+7, y = vOffset+4, labels = "Subcortical area", col="grey",xpd = T)
  brackets(151, vOffset+3, 164, vOffset+3,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")

for (jjj in 1:4){
for (iii in 1:length(unique(brainRegion))){
text(x = mean(which(brainRegion==unique(brainRegion)[iii])) + (jjj-1)*34, y = vOffset+1, labels = unique(brainRegion)[iii], col="grey",xpd = T)
  brackets(min(which(brainRegion==unique(brainRegion)[iii]))+ (jjj-1)*34, vOffset, max(which(brainRegion==unique(brainRegion)[iii]))+ (jjj-1)*34, vOffset,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
#text(x = mean(which(brainRegion==unique(brainRegion)[iii])), y = 48, '{', srt = -90, cex = 8, family = 'Helvetica Neue UltraLight')
}}
  for (iii in 1:length(unique(formatedPhe$category))){
  rect(xleft = 0.5, ytop = dim(allvar0)[1]+1.5-min(which(formatedPhe$category==unique(formatedPhe$category)[iii])), ybottom = dim(allvar0)[1]+0.5-max(which(formatedPhe$category==unique(formatedPhe$category)[iii])), xright = 164.5 , lwd = 2 )
}
  text(x = 154-14, y = vOffset+1, labels = "Left", col="grey",xpd = T)
  brackets(151-14, vOffset, 157-14, vOffset,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
  text(x = 161-14, y = vOffset+1, labels = "Right", col="grey",xpd = T)
  brackets(158-14, vOffset, 164-14, vOffset,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")

text(x = 154, y = vOffset+1, labels = "Left", col="grey",xpd = T)
  brackets(151, vOffset, 157, vOffset,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
  text(x = 161, y = vOffset+1, labels = "Right", col="grey",xpd = T)
  brackets(158, vOffset, 164, vOffset,h = 0.5,  type = 1, lwd = 2, lty = 1, xpd = T, col="grey")
dev.off()
```

## Write full table

```{r, message=FALSE, eval=F}

allvar02Write=allvar0
# Format and write table of results
for (jjj in 1:length(allvar0[1,])){
  for (iii in 1:length(allvar0[,1])){
    allvar02Write[iii,jjj]=paste0( signif(allvar0[iii,jjj],3), " (p=", formatC(allp0[iii,jjj], format = "e", digits = 1), ")")
  }
}
write.csv(allvar02Write, "STX_UKB_allROIS_Bsln_bdsz_covariates.csv")

```



<br><br>
