---
title: "Bivariate analyses to estimate grey-matter correlation"
author: "by [Baptiste Couvy-Duchesne] - `r format(Sys.time(), '%d %B %Y')`"
output:
  epuRate::PCTG:
    toc: TRUE
    code_folding: "show"
---

# Open phenotype data

I use read.csv here, which is a slower than *readr::read_csv()*. This is because I found that using *read_csv* opens the table as a *data frame* (not a *data table*) which does not allow to call TT[,columnName]

```{R, message=FALSE, eval=F}

TT=read.csv("UKB_phenotypes_15K_allData_regVars_Jan2019Update.csv")

```

# Write all pairwise phenotypes 

The aim of this code section is to write bivariate phenotype files that have the form **eid eid pheno1 pheno2** which will be read in OSCA to perform bivariate analyses.  

We restrict ourselves to phenotypes with significant morphometricity.  

```{R, message=FALSE, eval=F}

# Phenotype names
phe=read.csv("UKB_allvars_categories_fancylabels_withOldNames_Jan19Update.csv", stringsAsFactors = F)

# List of significant variables (correcting from baseline and body size variables)
signi=read.table("UKB_Phenotypes_signif_Bslne_bdsz.txt")

# Subset and get rid of covariates, for which bivariate models do not make sense
phe=phe[which(phe$V4 %in% signi$V1),]
phe=phe[-which(phe$V4 %in% c("Age_MRI", "weight_f21002_2_0",  "body_mass_index_bmi_f21001_2_0", "standing_height_f50_2_0", "sexuses_datacoding_9_f31_0_0")),]

# Loop and write files of phenotypes pairs
# We also store the list of of pairs of phenotypes that correspond to file names to use in the loop for parallel computing
var2write<-phe$V4

bivarList=NULL
# Bivariate files ICV regressed
for (iii in 1:length(var2write)){
  print(iii)
  for (jjj in (iii+1):length(var2write)){
    var1<-as.character(var2write[iii])
    var2<-as.character(var2write[jjj])
    write.table(TT[,c("eid", "eid", paste0(var1, "_regBB"), paste0(var2, "_regBB"))],paste0("UKB_phenotypes_bivariate_reg_Bsln_bdsz/",var1, "_", var2, ".txt" ), sep="\t", row.names=F, col.names=F , quote=F)
bivarList<-rbind(bivarList, paste0(var1, "_", var2))

  }
}

length(bivarList) # 53*52/2 = 1378 models to run
# Write list of pairwise combinations for loop
write.table(bivarList, "UKB_phenotypesPairs_bivariate_Bsln_bdsz.txt", col.names = F, row.names = F, quote = F)

```

# Run bivariate models in GCTA

**As bivariate analysis is not yet implemented in OSCA, we use GCTA. The trick is to rename the BRMs so GCTA sees them as GRM (genetic relatedness matrices).**  
GCTA also allows for multi-tread for faster computation (here 5 cpus per job).

```{bash, message=FALSE, eval=F}

od="working/directory/BREML_bivariate_reg_Bsln_bdsz"

cd ${wd}/BodFiles/
cp allVerticesBRM_QC.orm.bin allVerticesBRM_QC.grm.bin
cp allVerticesBRM_QC.orm.id allVerticesBRM_QC.grm.id
cp allVerticesBRM_QC.orm.N.bin allVerticesBRM_QC.grm.N.bin

cd ${od}
${bind}/qsubshcom "echo \${TASK_ID}.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_phenotypesPairs_bivariate_Bsln_bdsz.txt) |;  "${bind}"/gcta64 --reml-bivar 1 2 --grm-bin "${wd}"/BodFiles/allVerticesBRM_QC --pheno  "${wd}"/Phenotypes15K/UKB_phenotypes_bivariate_reg_Bsln_bdsz/\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_phenotypesPairs_bivariate_Bsln_bdsz.txt).txt --out  "${od}"/BivModel.\$(sed -n \"\${TASK_ID}{p;q}\" "${wd}"/UKB_phenotypesPairs_bivariate_Bsln_bdsz.txt).outliers.UKB --reml-maxit 10000 --reml-bivar-lrt-rg 0 --thread-num 5 " 5 20G GREML_all 48:00:00 "-array=1-1378 -acct=UQ-IMB-CNSG" 

```

# Extract rGM and phenotypic correlations

```{r, message=FALSE, eval=F}

# Rename columns to make code more readable
colnames(phe)=c("old_var" ,"category", "variable_clean", "my_var")

# Initialise - rGM matrices
rB<-matrix(data = NA, ncol =length(phe$my_var), nrow = length(phe$my_var) )
colnames(rB)<-phe$variable_clean
rownames(rB)<-phe$variable_clean
rBpval<-rB
rBSE<-rB

# Initialise - Phenotypic correlations
rP<-rB
rPpval<-rB
rPSE<-rB

# We need to store B1 and B2 (variance components) to calculate the residual correlation
BVC=matrix(data = NA, nrow =length(phe$my_var), ncol = 1 )
rownames(BVC)<-phe$variable_clean

# Loop on variable pairs and store values of interest
for (iii in 1:(length(phe$my_var)-1) ){
  for (jjj in (iii+1):length(phe$my_var)){

     var1<-phe$my_var[iii]
     var2<-phe$my_var[jjj]
     # rGM
     if(file.exists(paste0("BREML_bivariate_reg_Bsln_bdsz/BivModel.",var1, "_", var2, ".outliers.UKB.hsq"))){
out<-read.table(paste0("BREML_bivariate_reg_Bsln_bdsz/BivModel.",var1, "_", var2, ".outliers.UKB.hsq"), fill=T, stringsAsFactors = F)
rB[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[12,2]
rBpval[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[19,2]
 rBSE[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[12,3]  
     } else {
      if(file.exists(paste0("BREML_bivariate_reg_Bsln_bdsz/BivModel.",var2, "_", var1, ".outliers.UKB.hsq"))){
  out<-read.table(paste0("BREML_bivariate_reg_Bsln_bdsz/BivModel.",var2, "_", var1, ".outliers.UKB.hsq"), fill=T, stringsAsFactors = F)
rB[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[12,2]
rBpval[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[19,2]
 rBSE[phe$variable_clean[iii], phe$variable_clean[jjj]]<-out[12,3]     
       } }
     
# Get phenotypic correlation from written phenotype files too
if(file.exists(paste0("UKB_phenotypes_bivariate_reg_Bsln_bdsz/",var1, "_", var2, ".txt"))){
     dat=read.table(paste0("UKB_phenotypes_bivariate_reg_Bsln_bdsz/",var1, "_", var2, ".txt"))
     ct=cor.test(dat$V3, dat$V4)
     rP[phe$variable_clean[iii], phe$variable_clean[jjj]]<-ct$estimate
  rPpval[phe$variable_clean[iii], phe$variable_clean[jjj]]<-ct$p.value
   rPSE[phe$variable_clean[iii], phe$variable_clean[jjj]]<-sqrt( (1-ct$estimate**2) / ct$parameter )
} else {
  if(file.exists(paste0("UKB_phenotypes_bivariate_reg_Bsln_bdsz/",var2, "_", var1, ".txt"))){
    dat=read.table(paste0("UKB_phenotypes_bivariate_reg_Bsln_bdsz/",var2, "_", var1, ".txt"))
     ct=cor.test(dat$V3, dat$V4)
  rP[phe$variable_clean[iii], phe$variable_clean[jjj]]<-ct$estimate
  rPpval[phe$variable_clean[iii], phe$variable_clean[jjj]]<-ct$p.value
  rPSE[phe$variable_clean[iii], phe$variable_clean[jjj]]<-sqrt( (1-ct$estimate**2) / ct$parameter )
}
  
}
  }
  BVC[phe$variable_clean[iii]]=as.numeric(out[10,2])
}

# Write matrices and initialise plot
write.table(rB, "rB_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)
write.table(rBpval, "rB_pvalues_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)
write.table(rBSE, "rB_SE_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)

write.table(rP, "rP_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)
write.table(rPpval, "rP_pvalues_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)
write.table(rPSE, "rP_SE_correlation_matrix_outliers_Bsln_bdsz.txt", col.names=T, row.names = F)

```

# Calculate residual correlations

*findVCOVmatFromGCTAOutput()* is a R function that extract the matrix of variance covariace of model estimates from *qsubshcom* log files.  

```{r, message=FALSE, eval=F}

# Load functions to find matrix of var-covar of estimates and calculate the rE and SE 
source("RR_7.0_rE_SE_functions.R")

# Calculate residual correlations
rE<-matrix(data = NA, ncol =length(phe$my_var), nrow = length(phe$my_var) )
colnames(rE)<-phe$my_var
rownames(rE)<-phe$my_var
rESE<-rE
rEp<-rE

# Loop and calculate residual correlation and its SE
for (iii in 1:(length(phe$my_var)-1) ){
  print(iii)
  for (jjj in (iii+1):length(phe$my_var)){
    var1<-phe$my_var[iii]
     var2<-phe$my_var[jjj]
     infoo=findVCOVmatFromGCTAOutput(var1, var2, path = "BREML_bivariate_reg_Bsln_bdsz/job_reports/") 
     if (!is.null(infoo)){
  rE[var1,var2]=infoo[6,8]/sqrt(infoo[5,8]*infoo[4,8])
  rESE[var1,var2]=seRgFullApprox(VCX = infoo[5,8], VCY = infoo[4,8], CovXY = infoo[6,8], Var_CovXY = infoo[6,6], Var_VCX = infoo[4,4], Var_VCY = infoo[5,5], Cov_CovXY_VCX = infoo[4,6], Cov_CovXY_VCY = infoo[5,6], Cov_VCX_VCY = infoo[4,5])
     }
    }
}

# Calculate p-value
for (iii in 1:(length(phe$my_var)-1) ){
  print(iii)
  for (jjj in (iii+1):length(phe$my_var)){
    var1<-phe$my_var[iii]
     var2<-phe$my_var[jjj]
rEp[var1,var2]= (1 - pchisq( (rE[var1,var2] /rESE[var1,var2] )**2 , df = 1))/2
    }
}

write.table(rE, "rE_correlation_matrix_outliers_Bsln_bdsz_15K.txt", col.names=T, row.names = F)
write.table(rEp, "rE_pvalues_correlation_matrix_outliers_Bsln_bdsz_15K.txt", col.names=T, row.names = F)
write.table(rESE, "rE_SE_correlation_matrix_outliers_Bsln_bdsz_15K.txt", col.names=T, row.names = F)

```

# Plot grey-matter and residual correlations

```{r, message=FALSE, eval=F}

rB<-read.table("rB_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())
rBpval<-read.table("rB_pvalues_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())
rBSE<-read.table("rB_SE_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())

rp=read.table("rE_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())
rpp=read.table("rE_pvalues_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())
rpSE=read.table("rE_SE_correlation_matrix_outliers_Bsln_bdsz_15K.txt", header = T, colClasses = numeric())

rp=t(rp)
rpp=t(rpp)
rpSE=t(rpSE)

# Combine the two triangular matrics
allr<-matrix(0, nrow = length(phe$variable_clean), ncol = length(phe$variable_clean) )
allr[upper.tri(allr)]<-rB[upper.tri(rB)]
allr[lower.tri(allr)]<-rp[lower.tri(rp)]
colnames(allr)<-phe$variable_clean
rownames(allr)<-phe$variable_clean

allp<-matrix(0, nrow = length(phe$variable_clean), ncol = length(phe$variable_clean) )
allp[upper.tri(allp)]<-rBpval[upper.tri(rBpval)]
allp[lower.tri(allp)]<-rpp[lower.tri(rpp)]
allr[which(is.na(allr), arr.ind = T)]<-0

allSE<-matrix(0, nrow = length(phe$variable_clean), ncol = length(phe$variable_clean) )
allSE[upper.tri(allSE)]<-rBSE[upper.tri(rBSE)]
allSE[lower.tri(allSE)]<-rpSE[lower.tri(rpSE)]
allSE[which(is.na(allSE), arr.ind = T)]<-0

# Create object to highlight categories
ncate=as.data.frame(table(phe$category))
ncate$Freq
diag(allr)=NA

# Bonferroni significance threshold
signifUKB_Biv=0.05/(53*52)

# Make plot with correct colours and stars
png("Corplot_rE_rB_allstars_ordered_Bsln_bdsz_15K_stars.png", width = 30, height = 30, units = "cm", res = 300)
corrplot(corr = allr, method = "color",addgrid.col = "grey", type="full",  tl.cex = 1 ,  tl.col = "black", insig = "blank", col=colorRampPalette(c("#0072B2","#56B4E9","white", "#F0E442","#D55E00"))(100), na.label="square", tl.srt=70, cl.cex = 1.3)
for (iii in 1:length(unique(ncate$Var1))){
  rect(xleft = 0.5 + sum(ncate$Freq[0:(iii-1)]), ytop = 0.5+sum(ncate$Freq)-sum(ncate$Freq[0:(iii-1)]), ybottom = 0.5+sum(ncate$Freq) - ncate$Freq[iii] -sum(ncate$Freq[0:(iii-1)]), xright = 0.5 + ncate$Freq[iii] +sum(ncate$Freq[0:(iii-1)]), lwd = 3 )
}
stars=which(allp<signifUKB_Biv, arr.ind = T)
for (iii in 1:length(stars[,1])){
  text( stars[iii,2], dim(allr)[1] +1 - stars[iii,1], labels = "*", cex=1.2)
}
dev.off()


# write results
allrp=allr
for (iii in 1:dim(allr)[1]){
  for (jjj in 1:dim(allr)[1]){
allrp[iii,jjj]=paste0( signif(allr[iii,jjj], 3)," (SE=",signif(allSE[iii,jjj], 2) , "; p=", signif(allp[iii,jjj], 2), ")")
}
}
diag(allrp)=NA
write.csv(allrp, "UKB_Fig3a_rGM_rE_Bsln_bdsz.csv")

```

<br><br>
